---
title: "Final Project"
author: "Zhibo Hu"
date: "2019/4/25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(stringr)
library(ggplot2)
library(dplyr)
library(tidyr)
library(ggthemes)
library(rvest)
library(jsonlite)
library(RCurl)
library(caTools)
```

Get the raw data.
```{r,echo= FALSE}
#Get the raw data
FIFA19<- read_csv('https://raw.githubusercontent.com/amanthedorkknight/fifa18-all-player-statistics/master/2019/data.csv')
#Checking and cleaning the data structure
#str(FIFA19)
FIFA19 <- FIFA19[,-which(names(FIFA19) %in% c("X1","Photo","Flag","Club Logo","Real Face","Loaned From"))]
FIFA19 
```

Plot the Overall rating to see the distribution:
```{r,echo= FALSE}
#Plot the Overall rating to see the distribution
FIFA19 %>% ggplot(aes(Overall)) + geom_histogram(color = "white", fill = "black") +
  ggtitle("Player Overall rating") 
mean(as.numeric(FIFA19$Overall))
sd(as.numeric(FIFA19$Overall))
```
We can see that the Overall rating follows a normal distribution roughly with a mean of 66.2387 and a standard deviation of 6.90893.

FIFA19 has more than 20 different positions which are too trivial to analyze. Thus, I classify them into four groups: GK(Goal keeper); D(Defender); M(Middle); F(Forward).
Also, classify player age with 5 years intervel. Put players under 21 and player over 35 into to two big categories.
```{r,echo= FALSE}
#FIFA19 has more than 20 different positions which are too trivial to analyze. Thus, I classify them into four groups: GK(Goal keeper); D(Defender); M(Middle); F(Forward).
#Also, classify player age with 5 years intervel. Put players under 21 and player over 35 into to two big categories.
#Defind position groups.
pos <- unique(FIFA19$Position)
pos
D <- pos[str_detect(pos,"B$")]
F1 <- pos[str_detect(pos,"F$")]
F2 <- pos[str_detect(pos,"S$")]
F3 <- pos[str_detect(pos,"W$")]
F4 <- pos[str_detect(pos,"T$")]
F <- c(F1,F2,F3,F4)
M <- pos[str_detect(pos,"M$")]
D <- pos[str_detect(pos,"B$")]
GK <- "GK"

FIFA19 <- FIFA19 %>% 
  mutate(Position_group = ifelse(Position %in% GK, "GK", ifelse(Position %in% D, "D", ifelse(Position %in% M, "M", ifelse(Position %in% F, "F", "Unknown"))))) %>%
  mutate(Age_group = ifelse(Age <= 20, "Under 20", ifelse(Age > 20 & Age <=25, "21 to 25", ifelse(Age > 25 & Age <= 30, "25 to 30", ifelse(Age > 30 & Age <= 35, "31 to 35", "Over 35")))))
FIFA19

```

Plot1: Find the trend of overall rating with age growing up.
Plot2: Find the trend of overall rating with different positions.
```{r,echo= FALSE}
#Find the trend of overall rating with age growing up.
FIFA19 %>%
  filter(!Position_group %in% c("GK", "Unknown")) %>%
  group_by(Age) %>%
  summarise(Rating = mean(Overall)) %>%
  ggplot(aes(x= Age, y= Rating, group = 1)) +
  geom_line(color = "grey50", size = 1) +
  ggtitle("Age vs. Overall rating") 
#Find the trend of overall rating with different positions.
FIFA19 %>%
  group_by(Position_group, Age) %>%
  summarise(Rating = mean(Overall)) %>%
  ggplot(aes(x= Age, y= Rating, group = Position_group)) +
  geom_line(size = 1, color = "grey50") +
  facet_wrap(~ Position_group, ncol = 1)
```

The overall ratings increases before 31 and meets a inevitable decline around 34. 26-34 are the golden ages of a player. I remove the position "GK" because Goal Keeper is a special position that may have a much longer career.
For each position, it appears that Goal Keepers have the longest career life and Forwards have the shortest one. The decline of Middles start earliest around 33.

Finding the best overall and potential players for each age group.
```{r,echo= FALSE}
#Finding the best overall rated players for each age group.
best_rating <- FIFA19 %>%
  group_by(Age_group) %>%
  filter(Overall==max(Overall)) %>%
  arrange(Age_group) %>%
  select(ID,Name,Age_group,Overall)
#Finding the best potential players for each age group.
best_potential <- FIFA19 %>%
  group_by(Age_group) %>%
  filter(Potential==max(Potential)) %>%
  arrange(Age_group) %>%
  select(ID,Name,Age_group,Potential)
best_rating
best_potential
```
We can see that the players whose potential are the highest in their age group are already the best player in their age group.


The players' value and wage of raw data are shown like "€226.5M". Use Regex to transform them into numeric.
```{r,echo= FALSE}
#The players' value and wage of raw data are shown like "€226.5M". Use Regex to transform them into numeric.
FIFA19 <- FIFA19 %>% 
  mutate(Multiplier1 = ifelse(str_detect(Value, "K"), 1000, ifelse(str_detect(Value, "M"), 1000000, 1))) %>%
  mutate(PlayerValue = as.numeric(str_extract(Value, "\\d*\\d*\\.*\\d+")) * Multiplier1 ) %>%
  mutate(Multiplier2 = ifelse(str_detect(Wage, "K"), 1000, ifelse(str_detect(Wage, "M"), 1000000, 1))) %>%
  mutate(PlayerWage = as.numeric(str_extract(Wage, "\\d*\\d*\\.*\\d+")) * Multiplier2)
FIFA19
```

Plot player valuations.
```{r,echo= FALSE}
#Plot player valuations
FIFA19 %>%
  ggplot(aes(PlayerValue)) +
  geom_histogram(color = "white") +
  ggtitle("Player Valuations")  +
#Show the ouliers.
  geom_text(data = subset(FIFA19, Name == "Neymar Jr"), aes(x= PlayerValue, y= 500, label=Name), color = "blue") +
  geom_text(data = subset(FIFA19, Name == "L. Messi"), aes(x= PlayerValue, y= 1000, label=Name), color = "green") +
  geom_text(data = subset(FIFA19, Name == "K. De Bruyne"), aes(x= PlayerValue, y= 200, label=Name), color = "red") +
  geom_text(data = subset(FIFA19, Name == "E. Hazard"), aes(x= PlayerValue, y= 800, label=Name), color = "black") +
  geom_text(data = subset(FIFA19, Name == "Cristiano Ronaldo"), aes(x= PlayerValue, y= 800, label=Name), color = "purple")
```

The plot shows that player valuations are heavily skewed, for there are some superstar such as Messi and Ronaldo that are the outliers.

Find the top10 most expensive players.
```{r,echo= FALSE}
#Finding top 10 players with highest PlayerValue
FIFA19 %>%
  select(Name,PlayerValue,Club) %>%
  group_by(Club) %>%
  filter(PlayerValue==max(PlayerValue)) %>%
  arrange(desc(PlayerValue)) %>%
  head(10) %>%
  ggplot(aes(x=Name,y=PlayerValue,fill=factor(Club)))+geom_bar(stat='identity') + 
  labs(x='Player Name',y='PlayerValue',title='Top 10 most valuable players')
```


Plot valuations against positions.
```{r,echo= FALSE}
#Plot valuations against positions.
FIFA19 %>%
  filter(Position_group != "Unknown") %>%
  ggplot(aes(x= Position_group, y= PlayerValue)) +
  geom_boxplot() +
  scale_y_log10() +
  ggtitle("Player values vs. Positions") +
  xlab("Position")
#Specific positions.
FIFA19 %>%
  filter(Position_group != "Unknown") %>%
  ggplot(aes(x= Position, y= PlayerValue)) +
  geom_boxplot() +
  scale_y_log10() +
  coord_flip() +
  ggtitle("Player values vs. Specific Positions") +
  facet_wrap(~ Position_group, scales = "free")
```

Forwards and Middles are expensive than defenders and goal keepers.
For each specific position, left and right forwards are the most expensive positions.

Plot valuations against ages.
```{r,echo= FALSE}
#Plot valuations against ages.
FIFA19 %>%
  ggplot(aes(x= Age_group, y= PlayerValue)) +
  geom_boxplot() +
  scale_y_log10() +
  ggtitle("Players value vs. Age") +
  xlab("Age")
```

Players have the highest valuation in their 25-30 because of their golden age and longer expected career compared to the players in their 31-35. 

Plot valuations against overall rating.
```{r,echo= FALSE}
#Plot valuations against overall rating.
ggplot(FIFA19,aes(x=Overall,y=PlayerValue)) + geom_point(alpha=0.3, color = "blue") + 
  labs(x='Overall',y='Player Value',title='Player Value vs. Overall Rating')
```

The plot shows a positive relationship between player valuations and overall rating as expected.

Analyzing the mean rating for each of the attributes for each position except goal keeper.
Also, neglect the player whose overall rating is smaller than 75. 

```{r,echo= FALSE}
#Analyzing the mean rating for each of the attributes for each position except goal keeper.
p <- FIFA19 %>% select(ID, Position, Position_group)
#Create Attributes data.frame
Attibutes <- FIFA19 %>%
  filter(Position != "GK") %>%
  filter(Position != "Unknown") %>%
  filter(Overall >= 75) %>%
  select_if(is.numeric) %>% 
  select(-"International Reputation", -"Jersey Number", -"Skill Moves", -Special, -Multiplier1, -PlayerValue, -"Weak Foot", -Age, -Overall, -Potential, -Multiplier2, -PlayerWage) %>%
  left_join(p, by = "ID") %>%
  select(- ID) %>%
  gather(key = Attribute, value = Value, -Position, -Position_group) %>%
  group_by(Position_group, Position, Attribute) %>%
  summarise(mean = round(mean(Value, na.rm = T)))
Attibutes 
Attibutes %>%
  ggplot(aes(x= Attribute, y= Position)) +
  geom_tile(aes(fill = mean), colour = "black") +
  geom_text(aes(label = mean)) +
  scale_fill_gradient(low = "purple", high = "green") +
  theme(axis.text.x = element_text(angle = 40, hjust = 1), strip.text = element_text(face = "bold", size = 1), legend.position = "none") 
```

The results show that forwards and middles do well in agility, acceleration and skills such as dribbling and shots, while defenders are good at strength.

Check when does the overall rating and potential converge.
```{r,echo= FALSE}
#Plot the average overall line and the average potential line together.
FIFA19 %>%
  group_by(Age) %>%
  summarise(Potential = mean(Potential),Overall = mean(Overall)) %>%
  ggplot(aes(Age)) +
  geom_line(aes(y= Potential), color = "blue", size = 1) +
  geom_line(aes(y= Overall), color = "black", size = 1) +
  annotate("text", x= 30, y=71, label = "Converge point", color = "red") +
  ggtitle("Potential And Overall Talent against age")
```

Potential and Overall rating coverge at 30 years old, which means players reach highest overall rating at age 30. 

As a team manager, finding and developing potential younger player is a importance job. Give a Anlysis of the young players.
```{r,echo= FALSE}
#Defind growth=Potential - Overall and plot the young players with large growth.
Young <- FIFA19 %>% mutate(Grow = Potential - Overall) %>%
  arrange(desc(Grow)) %>%
  head(20) %>%
  select(Name, Value, Overall, Age, Grow, Position) %>%
  gather(key = Mode, value = ExpectedValue, -Name, -Value, -Age, -Position) %>%
  mutate(MOde = factor(x=Mode, levels = c("Grow", "Overall")))
Young
Young %>%
  ggplot(aes(Name, ExpectedValue, fill = Mode)) +
  geom_bar(stat = "identity", position = "stack", color = "black") +
  coord_flip() +
  scale_fill_manual(values = c("Overall" = "green", "Grow" = "purple")) +
  ggtitle("Player's Growth") +
  theme(legend.position = "none") +
  geom_text(data = subset(Young, Mode == "Overall"), aes(Name, 10, label=paste(Position, Value, Age, sep = ", "))) 
```

The green parts show the young players' overall rating right now and the purple parts are the space they have to reach their potential. It is wise for a manager to focus more on the young players that have higher potential and more space to grow and also have a low market value. Players like B.Mumba and Y.Lenze are good bargain.

See the age distribution of the highest value clubs.
```{r,echo= FALSE}
#Find the clubs with highest valuation. (sum of playervalue)
ClubValue <- FIFA19 %>%
  group_by(Club)%>%
  summarise(ClubValue = round(sum(PlayerValue)/1000000))%>%
  arrange(desc(ClubValue)) %>%
  head(10)
ClubValue
#Use club's name to find players
PlayerList <- list()
for (i in 1:nrow(ClubValue)){
name <-  FIFA19 %>%
  filter(Club == ClubValue$Club[i]) 
PlayerList[[i]] <- name
}
a <- lapply(PlayerList, as.data.frame) %>% bind_rows()
a$Club <- as.factor(a$Club)
ggplot(a, aes(x = Club ,y = Age)) +
  geom_boxplot()+
  theme(axis.text.x = element_text(angle = 90), legend.position = "none")+
  coord_flip() +
  ylab("Age distribution")
```
Hotspur seems to have a wonderful balance between players age and valuations. Juventus's players are older than other teams which means they may need big changes in the future to find more young players.

Find the top 20 teams with highest average overall rating.
```{r,echo= FALSE}
#Find the top 20 teams with highest average overall rating.
Overall_team <- FIFA19 %>%
  group_by(Club) %>%
  summarise(Average_rating = mean(Overall, na.rm = T)) %>%
  arrange(desc(Average_rating)) %>%
  head(20) %>% 
  pull(Club) 
Overall_team

FIFA19 %>%
  filter(Club %in% Overall_team) %>%
  mutate(Italy_club = ifelse(Club %in% c("Juventus", "Napoli", "Inter","Milan","Roma","Lazio"), "Yes", "No")) %>%
  ggplot(aes(x= reorder(Club,Overall), y= Overall, fill = Italy_club)) +
  geom_boxplot(color = "black") +
  scale_fill_manual(values = c("grey50", "purple")) +
  ggtitle("Top 20 Highest Overall Ratings") +
  coord_flip() +
  theme(legend.position = "none")
```

Listed above are the 20 best teams with highest average rating. They are the strongest football club in to world. There are 6 italian clubs in the rank and 4 of them are in the top 5.

Define superstar as a player whose rating is equal or higher than 85. Find the teams that have most superstars.
```{r,echo= FALSE}
#Define superstar as a player whose rating is equal or higher than 85. Find the teams that have most superstars.
FIFA19 %>%
  mutate(superstars = ifelse(Overall >= 85, "Yes", "Not")) %>%
  group_by(Club, superstars) %>%
  filter(superstars == "Yes") %>%
  summarise(NumberofSuperstar = n()) %>%
  filter(NumberofSuperstar >1) %>%
  mutate(Italian = ifelse(Club %in% c("Juventus", "Napoli", "Inter","Milan","Roma","Lazio"), "Yes", "No")) %>%
  arrange(desc(NumberofSuperstar)) %>%
  ggplot(aes(x= reorder(Club,NumberofSuperstar), y= NumberofSuperstar, fill = Italian)) +
  geom_col(color = "black") +
  scale_fill_manual(values = c("grey50", "purple")) +
  ggtitle("Number of Superstars") +
  scale_y_continuous(breaks = seq(0,12,1))+
  coord_flip() +
  theme(legend.position = "none")
```

Real Madrid, Manchester City and Juventus are the top3 club that have most superstar. Some Italian teams like Inter, Roma and Lazio with higher average rating only have a few superstars. This plot may better show the real power of a club because only 14 players can play in one game. Superstars represent the real talent of the club.

Find the teams with highest attacking and defending abilities. Defind attacking with the average of Finishing, Longshots, ShotPower and Positioning，Dribbling and BallContro. Defend defending with average of
Interceptions,Aggression,Jumping and Strength.
```{r,echo= FALSE}
#Defind attacking and find the best teams.
FIFA19  %>%
  mutate(Attacking = (Finishing + LongShots + ShotPower + Positioning + Dribbling + BallControl) / 6) %>%
  filter(Position_group %in% c("M", "F")) %>%
  group_by(Club) %>%
  summarise(NumberofAttackers = n(), TeamAttacking = mean(Attacking)) %>%
  arrange(desc(TeamAttacking)) %>% head(20) 
#Defind defending and find the best teams.
FIFA19  %>%
  mutate(Defending = (Interceptions + Aggression + Jumping + Strength) / 4) %>%
  filter(Position_group %in% c("D")) %>%
  group_by(Club) %>%
  summarise(NumberofDefender = n(), TeamDefending = mean(Defending)) %>%
  arrange(desc(TeamDefending)) %>% head(20) 
```
As expected, most of the best attacking and defending clubs are also the clubs with highest average overall rating and most superstars. This statistic shows that soccer is a game with both attacking and defending. Only focus on them both can make the club strong and powerful.

Team wage analysis
First, find the top10 players who have the highest wages
```{r,echo= FALSE}
#Find the top10 players who have the highest wages
FIFA19 %>%
  select(Name,PlayerWage,Club) %>%
  group_by(Name) %>%
  filter(PlayerWage == max(PlayerWage)) %>%
  arrange(desc(PlayerWage)) %>%
  head(10) %>%
  ggplot(aes(x=Name,y=PlayerWage,fill=Club)) + geom_bar(stat = 'identity') +
  labs(x='Player Name',y='Wages drawn by players',title='Top 10 players with highest wages')
```

```{r,echo= FALSE}
FIFA19 %>% 
  group_by(Club) %>% 
  summarise(TotalWages = sum(PlayerWage, na.rm = TRUE),
            Average_rating = mean(Overall, na.rm = TRUE)) %>%
  mutate(ValueForMoney = TotalWages / Average_rating) %>%
  arrange(desc(TotalWages)) %>% head(20) %>%
  ggplot(aes(x= reorder(Club, TotalWages), y= TotalWages)) +
  geom_col(colour = "white") +
  geom_text(aes(label = round(ValueForMoney), hjust = 0)) +
  scale_y_continuous() +
  coord_flip() +
  ggtitle("Wages") +
  theme(legend.position = "none") +
  xlab("Team") + ylab("Wage amounts")
```

The number after the bar shows how much a club pays for one point of rating. We can find that Spanish clubs such Real Madrid and FC Barcelona spend a lot of money to pay the players' wages in order to get great achievemnets.

BMI is a way to measure a player's feature. BMI larger than 25 is considered to be overweight. In order to compute BMI, I have to transform the height into inches and extract the number in weight using function and Regexp.
```{r,echo= FALSE}
#Write a function using Regexp to extract the number.
extract <- function(x){
  regexp <- "[[:digit:]]+"
  str_extract(x, regexp)
}
#Extract the number from weight
weight <- sapply(FIFA19$Weight, extract)
FIFA19$Weight <- as.numeric(weight)
#Transform height into inch。
split <- str_split(FIFA19$Height, "'")
for (i in 1:length(split)){
  split[[i]] <- as.numeric(split[[i]])
} 
for (i in 1:length(split)){
  split[[i]] <- (split[[i]][1] * 12 ) + split[[i]][2]
}
FIFA19$Height <- as.numeric(unlist(split))
FIFA19
```

Then, compute BMI. All the players are in a normal shape as expected for they are athletes. 
```{r,echo= FALSE}
#Compute BMI
FIFA19%>%
  group_by(Name)%>%
  mutate(BMI = (Weight*0.454/(Height)^2)*1000)%>%
  arrange(desc(BMI))%>%
  select(Name, BMI)%>%
  head(10)
```


The sofifa.com releases the hottest 100 player in the past week. Scraping the webpage to get player's raw data.
```{r,echo= FALSE}
#Scraping data
table_url <- "https://sofifa.com/players/hot"
Hot100 <- table_url %>%
  read_html() %>%
  html_node(xpath = "//*[@id='content-target']/table") %>%
  html_table()
Hot100 <- Hot100[,-c(1,7,10,12)]
```

Cleaning the raw data.
```{r,echo= FALSE}
#Set colnames
colnames(Hot100)=Hot100[1,]
Hot100<-as.data.frame(Hot100[-1,]) 
#Remove the useless position information in Name
Hot100[,1] <- gsub("[A-Z]{2,}","",Hot100$Name)
Hot100
```

Join the tables of Hottest player and FIFA19 to see their values and rating.
```{r,echo= FALSE}
#Join the tables of Hottest player and FIFA19
HotPlayer <- inner_join(FIFA19,Hot100, by = "Name") %>%
  select(-ID,-Multiplier1,-Multiplier2)
HotPlayer<-rename(HotPlayer,"Reputation"="International Reputation")
HotPlayer
```

Use cor function to find the factors that related to PlayerValue. Choose the factors whose corrlation with PlayValue is larger than 0.65.
```{r,echo= FALSE}
#Find the factors that related to PlayerValue.
Player_Value <- HotPlayer[ , map_lgl(HotPlayer, is.numeric)]
mcor<- as.data.frame(cor(Player_Value, use = "complete.obs"))
#Choose the factors whose corrlation with PlayValue is larger than 0.65.
factor <- rownames(subset(mcor["PlayerValue"], PlayerValue > 0.65))
factor
```

Use the HotPlayer table and the factors found above to fit a linear regression using PlayerValue as response.
```{r,echo= FALSE}
#Fit a linear model. Split the data into train and test.
set.seed(101)
sample = sample.split(HotPlayer, SplitRatio = 0.6)
train <- subset(HotPlayer, sample == TRUE)
test <- subset(HotPlayer, sample == FALSE)
lm <- lm(PlayerValue ~ Overall + Reactions + Special + Reputation + PlayerWage, data = train, na.action = na.omit)
summary(lm)
```

The model's R^2 is larger than 0.9 and most of the estimators are significantly larger than 0.
Then, use the model to predict players' value in test data. Defind prediction accurracy if the difference between prediction and true value is smaller than 1/5 of the prediction.

```{r,echo= FALSE}
test_lm <- predict(lm, newdata = test)
test$Predicted.Value <- test_lm
value_predict <- test[c("Name","PlayerValue","Predicted.Value")]  %>%
  mutate(Difference = PlayerValue - Predicted.Value )
value_predict$Accuracy <- ifelse(value_predict$Difference > 0.20 * value_predict$PlayerValue , "No",ifelse(value_predict$Difference < -(0.20 * value_predict$PlayerValue),"No", "Yes"))
value_predict%>%
  head(10)
```

4 in 10 predictions give an accurate predict. The model tends to give a larger value compare to the true value, which means that the model still needs to be improved.

```{r,echo= FALSE}

```



